---
output:
    github_document:
        html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```

# CAMDAC

Copy-number Aware Methylation Deconvolution Analysis of Cancer (CAMDAC) is an R library for deconvolving bulk tumor DNA methylation (bisulfite) sequencing data ([Larose Cadieux et al., 2022, bioRxiv](https://www.biorxiv.org/content/10.1101/2020.11.03.366252v2)).

<!-- badges: start -->
<!-- badges: end -->

## Documentation

Visit [https://vanloo-lab.github.io/CAMDAC/](https://vanloo-lab.github.io/CAMDAC/).

## Installation : Dockerhub

A CAMDAC container is available on [dockerhub](https://hub.docker.com/r/nmensah5/camdac) for use with Docker, Singularity or Apptainer:

```{bash}
docker pull nmensah5/camdac:latest
echo "library(CAMDAC)" > commands.R
docker run -v $(pwd):/app nmensah5/camdac:latest Rscript commands.R
```

## Installation : Github

You can install CAMDAC and its dependencies from an R console:

```{r}
install.packages("remotes")
remotes::install_github("VanLoo-lab/CAMDAC")
```

## Quickstart 

We provide pre-built reference datasets for hg38 and hg19. These files are required to run CAMDAC for either RRBS or WGBS analysis [from the Zenodo repository: (10565423)](https://zenodo.org/records/10565423). An R getter function is provided for convenience:

```{r}
CAMDAC::download_pipeline_files(bsseq = "rrbs", directory = "./refs")
CAMDAC::download_pipeline_files(bsseq = "wgbs", directory = "./refs")
```

For WGBS analysis, CAMDAC requires the `java` command line utility to be available in the system PATH.

With reference files downloaded, run the tumor-normal deconvolution pipeline with test data:

> [!NOTE]  
> We provide downsampled BAM files for testing the pipeline. For representative results, please use your own BAM files.

```{r example}
library(CAMDAC)

tumor_bam <- system.file("testdata", "tumour_beds_min.sorted.bam", package = "CAMDAC")
normal_bam <- system.file("testdata", "normal_beds_min.sorted.bam", package = "CAMDAC")

# Select samples for basic tumor-normal analysis
tumor <- CamSample(id = "T", sex = "XY", bam = tumor_bam, patient_id="readme")
normal <- CamSample(id = "N", sex = "XY", bam = normal_bam, patient_id="readme")

# Configure pipeline
config <- CamConfig(
  outdir = "./validation/results/test_readme/", bsseq = "rrbs", lib = "pe",
  build = "hg38", refs = "./refs", n_cores = 1, cna_caller='ascat',
  min_cov=1, # Minimum tumour coverage at 1 for testing.
  min_normal_cov=1, # Minimum normal coverage at 1 for testing. 
  min_mapq=1 # Minimum MAPQ at 1 for testing.
)

# Run CAMDAC
CAMDAC::pipeline(
  tumor, germline = normal, infiltrates = normal, origin = normal, config
)
```

For a more detailed walkthrough with test data, see `vignette("pipeline")`.

## Contributing

To contribute to CAMDAC, fork [the repository](https://github.com/VanLoo-lab/CAMDAC) and install the development dependencies with `remotes::install_dev_deps('.')`.

After making your changes, run the build and test commands listed in `vignette("contributing")`.

Finally, submit a [pull request](https://github.com/VanLoo-lab/CAMDAC/pulls) with the changes on your fork.
