---
title: "allele-specific"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{allele-specific}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

##  Allele-specific methylation (ASM) analysis

CAMDAC can be used to detect allele-specific methylation (ASM) by phasing CpGs to heterozygous SNPs and deconvolving bulk methylation rates per allele. 

This tutorial steps through the ASM analysis pipeline. Briefly:
- A copy number solution is generated for the tumor sample if not available
- Bulk allele-specific methylation rates are inferred on samples using SNP loci
- CpG copy number is inferred from BAF rates
- Allele-specific methylation rates are deconvolved from the tumor sample
- Differential ASM is called within and between samples

## CAMDAC-ASM from BAM files

CAMDAC ASM analysis is run using the CAMDAC `asm_pipeline()` function. This pipeline generates the allele-specific copy number solution and heterozygous SNP loci, before deconvolution and differential ASM analysis:

```{r}
b_tumor <- system.file("testdata", "tumor.bam", package = "CAMDAC")
b_normal <- system.file("testdata", "normal.bam", package = "CAMDAC")

tumor <- CamSample(id="T", sex="XY", bam=b_tumor)
normal <- CamSample(id="N", sex="XY", bam=b_normal)
config <- CamConfig(outdir="./results", bsseq="wgbs", lib="pe", build="hg38", n_cores=10)

asm_pipeline(
    tumor=tumor,
    germline=normal,
    infiltrates=normal,
    origin=normal,
    config=config
)
```

The ASM results directory contains the following files. [TODO]

## TODO: CAMDAC-ASM from external inputs

To run the ASM pipeline without BAM files, CAMDAC requires:
- Each CamSample object has SNP loci
- The tumor CamSample object has an allele-specific CNA solution
- All CamSample objects have BAM files available for phasing

CAMDAC-ASM requires a file of heterozygous SNP loci against which CpGs will be phased. This is a tab-delimited file with a header containing four fields:

| Field | Description |
| ---- | ------------ |
| chrom | Chromosome name |
| pos | SNP loci position |
| ref | The reference allele (A/C/T/G) |
| alt | The alternate SNP allele (A/C/T/G) | 

## ASM analysis after CAMDAC main pipeline 

TODO

## ASM analysis with external inputs

First, attach your SNP loci file to the tumor object with `attach_output()`, then run `asm_pipeline()`:
```{r}
# Setup CAMDAC samples
tumor <- CamSample(id="tumor", sex="XY", bam=b_tumor)
normal <- CamSample(id="normal", sex="XY", bam=b_normal)
config <- CamConfig(outdir="./results", ref="./pipeline_files", bsseq="wgbs", lib="pe", cores=10)

# Add SNPs
asm_snps_file <- system.file("testdata", "test_het_snps.tsv", package = "CAMDAC")
attach_output(tumor, config, "asm_snps", asm_snps_file)
attach_output(normal, config, "asm_snps", asm_snps_file)
```

Next, CAMDAC requires the allele-specific copy number solution from the tumor, attached as follows:
```{r}
cna_file <- system.file("testdata", "test_cna.tsv", package = "CAMDAC")
attach_output(tumor, config, "cna", cna_file)
```

Finally, run the allele-specific methylation pipeline:
```{r}
asm_pipeline(
  tumor=tumor,
  infiltrates=normal,
  origin=normal,
  config=config
)
```

Results from this pipeline are found in the results directory under 'AlleleSpecific'. `The pipeline analyses data with the following steps:

1. Count CpG methylation on tumor and normal at sites phased to SNP loci.
2. Deconvolve methylation on tumor **per haplotype** using the normal
3. Assign allele-specific copy number state **per CpG** using the bulk tumor solution
4. Call allele-specific differential methylation within samples
5. Call allele-specific differential methylation between samples

## Using SNPs called from previous CAMDAC runs

If you have already run the CAMDAC pipeline in tumor-normal mode, then the tumor object's SNP files will be used by default.
