---
title: "CAMDAC pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pipeline}
  %\VignetteEngine{rmarkdown::render}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

The entry-point to CAMDAC is the `pipeline()` function which expects a `CamConfig()` object and four `CamSample()` objects representing:

* **tumor** : The bulk tumor sample to deconvolve
* **germline** : The germline normal data for copy number calling
* **infiltrates** : A proxy for the normal infiltrating cells
* **origin** : A proxy for the normal cell from which the tumour originated

The same normal sample may be passed repeatedly for the germline, infiltrates or origin, depending on your experimental design. See `?pipeline` for more details.

```{r pipeline}
library(CAMDAC)

# Path to BAM files
tumor_bam <- system.file("testdata", "tumor.bam", package = "CAMDAC")
normal_bam <- system.file("testdata", "normal.bam", package = "CAMDAC")

# Select samples for basic tumor-normal analysis
tumor <- CamSample(id = "T", sex = "XY", bam = tumor_bam)
normal <- CamSample(id = "N", sex = "XY", bam = normal_bam)

# Configure pipeline
config <- CamConfig(
  outdir = "./results", bsseq = "rrbs", lib = "pe",
  build = "hg38", refs = "./refs", n_cores = 1, cna_caller = 'ascat'
)

# Run CAMDAC
CAMDAC::pipeline(
  tumor, germline = normal, infiltrates = normal, origin = normal, config
)
```

Next, see `vignette("output")` for a detailed summary of CAMDAC results files.
