---
title: "Installation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{setup}
  %\VignetteEngine{rmarkdown::render}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Install CAMDAC

From the R console, install CAMDAC from github:

```{r}
install.packages("remotes")
remotes::install_github("VanLoo-lab/CAMDAC")
```

## Download pipeline reference files

CAMDAC requires custom annotation files for RRBS and WGBS analysis, available [at the Zenodo repository: (10565423)](https://zenodo.org/records/10565423). An R convenience function is provided to download these files:

```{r}
CAMDAC::download_pipeline_files(bsseq = "rrbs", directory = "./refs")
CAMDAC::download_pipeline_files(bsseq = "wgbs", directory = "./refs")
```

Now, you're ready to run CAMDAC! Next, see `vignette("pipeline")`.

### Reference file search priority

CAMDAC searches for pipeline files in the following order:

1. A directory passed when creating the config object (see `CamConfig()`)
1. The location defined by the environment variable CAMDAC_PIPELINE_FILES.
1. The current working directory

We recommend that you set the environment variable `CAMDAC_PIPELINE_FILES` to the directory where you downloaded the files. This will allow CAMDAC to find the files automatically whenever you load R.

From a Unix terminal:

> echo "CAMDAC_PIPELINE_FILES=$(realpath R)" >> ~/.Renviron

## External dependencies

**CAMDAC-RRBS**

* None

**CAMDAC WGBS**

* `java`:  To run CAMDAC on WGBS data, we leverage Battenberg which requires the `java` command-line utility. Download Java from [https://openjdk.org/](https://openjdk.org/).